---
title: "Day 8 Notes"
author: "Andre Dextre"
date: "2022-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(gapminder)
library(sf)
```

### Read in lizard data

```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

```{r}
ggplot(data = lizards) +
  geom_point(aes(x = total_length,
                           y = weight))
```

Create a histogram of total lengths for all lizards in the dataset.

```{r}
ggplot(data = lizards, aes(x = total_length)) +
  geom_histogram(color = "orange",
                 fill = "purple",
                 size = 1,
                 linetype = "dotted")
```

```{r}
ggplot(data = lizards, aes(x = total_length,
                           y = weight)) +
  geom_point(shape = 22,
             color = "cyan4",
             fill = "yellow",
             size = 4,
             alpha = 0.4
  )
```

```{r}
ggplot(data = lizards, aes(x = total_length,
                           y = weight)) +
  geom_point(aes(color = common_name,
                 size = total_length),
             alpha = 0.5) +
  theme_minimal()
```

```{r}
ggplot(data = lizards, aes(x = total_length,
                           y = weight)) +
  geom_point() +
  facet_wrap(~common_name, scales = "free")
```

```{r}
lizard_count <- lizards %>%
  group_by(common_name) %>%
  summarize(counts = n())

lizard_counts <- lizards %>%
  count(common_name)

ggplot(data = lizard_count, aes(y = common_name,
                                x = counts)) +
  geom_col()
```


```{r}
ggplot(data = lizard_count, aes(y = fct_reorder(common_name, counts),
                                x = counts)) +
  geom_col()
```

Create scatter plots, faceted by species, for total length vs weight. The color and shape of points should differ based on the species (common_name). Change the theme. Make each faceted panel have flexible axis limits (scales should be free).

```{r}
ggplot(data = lizards, aes(x = total_length,
                           y = weight)) +
  geom_point(aes(color = common_name),
             alpha = 0.6,
             size = 1) +
  theme_minimal() +
  labs(x = "Total Length (mm)",
       y = "Weight (g)",
       title = "Jornada Basin Lizard Sizes",
       subtitle = "All lizards follow the standard length-weight relationship, with Western Whiptails being largest",
       caption = "Data: Jornada Basin LTER, CCBY") +
  facet_wrap(~common_name, scales = "free")

```

## Types of Scales

Qualitative Scales: To distinguish between groups

Sequential Scale: To indicate values or value order

Diverging Scales: When there's an obvious "mid" point, and you want to show how much higher and lowers things are from it.

Highlights: To point out something of interest

```{r}
ggplot(data = lizards, aes(x = total_length,
                           y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_steps(low = "red",
                    high = "black")
```


```{r}
ggplot(data = lizards, aes(x = total_length,
                           y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_steps2(low = "purple",
                     mid = "white",
                    high = "orange",
                    midpoint = 150,
                    breaks = c(50, 75, 150, 180, 220, 280))
```


```{r}
ggplot(data = lizards, aes(x = total_length,
                           y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_stepsn(colors = c("orange", "red", "purple"),
                     breaks = seq(from = 0, to = 300, by = 60))
```

Make a horizontal box plot with common name on the y-axis, total length on the x-axis, with box plot fill changing based on the common name. 

```{r}

ggplot(lizards, aes(x = total_length,
                    y = common_name)) +
  geom_boxplot(aes(fill = common_name))

```


```{r}
lizards_median <- lizards %>%
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

ggplot(lizards_median, aes(x = total_length,
                           y = common_name)) +
  geom_boxplot(aes(fill = common_name))
```

```{r}

ggplot(lizards_median, aes(x = total_length,
                           y = common_name)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "ggsci::default_gsea") +
  theme_minimal()
```

```{r}
ggplot(lizards_median, aes(x = total_length,
                           y = common_name)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "ggsci::legacy_tron") +
  theme_minimal()
```


```{r}
ggplot(lizards, aes(x = total_length,
                    y = weight)) +
  geom_point(aes(color = weight)) +
  theme(panel.grid.major.x = element_line(color = "red"),
        panel.grid.minor.y = element_blank(),
        axis.title = element_text(color = "purple", size = 10),
        axis.text.y = element_text(color = "green"),
        text = element_text(size = 18),
        panel.background = element_rect(color = "purple", fill = "yellow")) +
  annotate("text", x = 300, y = 50, 
           label = "Woooooo", color = "blue") +
  geom_vline(xintercept = 250, linetype ="dashed",
             color = "orange", size = 4)
```

Repulsive labels with `ggrepel`

```{r}
wws_lizards <- lizards %>%
  filter(common_name == "western whiptail",
         site == "sand")

ggplot(wws_lizards, aes(x = total_length,
                        y = weight)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num))
```


```{r}
gapminder %>%
  filter(year == 2002,
         continent == "Europe") %>%
  ggplot(aes(x = gdpPercap,
             y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country), size = 3)

```

```{r}
ggplot(lizards, aes(x = total_length,
                        y = weight)) +
  geom_point() +
  gghighlight(weight > 30, label_key = toe_num)
```

### Make a Map!

Read in the spatial data 

```{r}
jornada_veg <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml"))

ggplot() +
  geom_sf(data = jornada_veg,
          aes(fill = Name),
          color = NA) +
  theme_minimal() +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys")
```

